---
title: "Matching With Replacement (Optional Bootstrap)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Matching With Replacement (Optional Bootstrap)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    eval = FALSE
)
```

## Overview

This vignette demonstrates how to run the CohortBuilder matching pipeline in
**with-replacement** mode, where each eligible exposed spell is matched to a
control spell by minimising a random-number difference via SQL inside DuckDB.
Optionally, the match can be repeated across `n_bootstraps` bootstrap resamples
to quantify variability.

## Prerequisites

```{r setup}
library(CohortBuilder)
library(data.table)
```

## Step 1 — Configure parameters

```{r parameters}
# Variables to match on (must be present in the eligibility data)
matching_vars <- c(
    "SV_SEX", "SV_REGION", "SV_HIST_COVID_VACC", "SV_PRIOR_COVID_DG",
    "SV_BRAND_COVID_VACC", "SV_PREG_STATUS", "SV_IMMUNOCOMPROMISED",
    "CDC_RISK", "SV_SES_STATUS"
)

# --- Directories --------------------------------------------------------------
dir_input <- "input"
dir_int_outputs <- "intermediate_outputs"
dir_matching_db <- "intermediate_outputs/matching.duckdb"
dir_bootstrap <- "intermediate_outputs/bootstraps"
dir_outputs <- "outputs"

# --- Matching options ----------------------------------------------------------
save_output <- FALSE
save_intermediate_matching_ouputs <- TRUE
with_bootstrap <- FALSE # set TRUE to run bootstrap resampling
n_bootstraps <- 5 # number of bootstrap iterations
start_seed <- 42
```

## Step 2 — Prepare input data

Provide your own eligibility dataset as a `data.frame` or `data.table` saved in
`input/D3_ELIGIBILITY.parquet`, or generate synthetic data for testing:

```{r input}
# Option A: generate synthetic data
D3_ELIGIBILITY <- generate_eligibility_data(
    n = 100000,
    start_seed = start_seed,
    save_output = FALSE,
    output_dir = dir_input,
    output_file = "D3_ELIGIBILITY"
)

# Option B: read your own data
# D3_ELIGIBILITY <- arrow::read_parquet(file.path(dir_input, "D3_ELIGIBILITY.parquet"))
```

The minimum required columns are:

| Column | Description |
|---|---|
| `person_id` | Unique person or spell identifier |
| Matching variables | As listed in `matching_vars` |
| `eligible_exposed` | Logical — TRUE if the row is an exposed spell |
| `eligible_control` | Logical — TRUE if the row is an eligible control |
| `matching_status_start` | Date: start of the matching window |
| `matching_status_end` | Date: end of the matching window |
| `year_of_birth` | Integer year of birth (used for ± 1-year age matching) |

## Step 3 — Load SQL templates

The package ships the SQL query templates. Resolve them with `system.file()`:

```{r sql}
matching_query <- suppressWarnings(
    getSQL(system.file("sql_queries", "matching_query_with_replacement.sql",
        package = "CohortBuilder"
    ))
)
target_table_query <- getSQL(
    system.file("sql_queries", "create_matching_target_table.sql",
        package = "CohortBuilder"
    )
)
```

## Step 4 — Run the pipeline

```{r run}
D4_MSC <- build_study_cohort(
    eligible_pop = D3_ELIGIBILITY,
    matching_query = matching_query,
    target_table_query = target_table_query,
    matching_vars = matching_vars,
    dir_matching_db = dir_matching_db,
    n_cores = NULL, # auto-detects all cores minus one
    log_name = "log_build_study_cohort",
    output_pars = list(
        save_output      = save_output,
        dir_output       = dir_outputs,
        output_file_name = "D4_MSC"
    ),
    intermediate_output_pars = list(
        save_intermediate_outputs = save_intermediate_matching_ouputs,
        dir_intermediate_outputs  = dir_int_outputs,
        profile_table_name        = "D3_LOOKUP_TABLE",
        matching_pop_name         = "D3_MATCHING_POP"
    ),
    bootstrap_pars = list(
        with_bootstrap = with_bootstrap,
        n_bootstraps   = n_bootstraps,
        dir_bootstrap  = dir_bootstrap,
        start_seed     = start_seed
    ),
    input_column_names = list(
        col_person_id = "person_id",
        col_eligible_exposed = "eligible_exposed",
        col_eligible_control = "eligible_control",
        col_matching_status_start = "start",
        col_matching_status_end = "end",
        col_age_iterator = "year_of_birth"
    ),
    output_column_names = list(
        col_person_id       = "person_id",
        col_match_id        = "match_id",
        col_treatment_group = "group",
        col_T0              = "T0"
    ),
    matching_mode = "with_replacement"
)
```

## Step 5 — Inspect results

When `save_output = FALSE` (or when returned), `D4_MSC` is a `data.table` with
one row per person-spell in the matched cohort:

```{r inspect}
# Group counts
D4_MSC[, .N, by = group]

# Number of matched pairs
D4_MSC[!is.na(match_id), data.table::uniqueN(match_id)]
```

| Column | Description |
|---|---|
| `person_id` | Person/spell identifier |
| `match_id` | Shared identifier for each matched pair |
| `boot_id` | Bootstrap iteration (always 1 when `with_bootstrap = FALSE`) |
| `group` | `"EXPOSED"`, `"CONTROL"`, or `"UNMATCHED"` |
| `T0` | Index date (start of the exposed spell) |
| `matching_status_start` | Start of the eligibility window |
| `matching_status_end` | End of the eligibility window |
| Matching variables | Copied from the lookup table |

## Bootstrap mode

Set `with_bootstrap = TRUE` and `n_bootstraps` to a positive integer. Each
bootstrap resample is written to `dir_bootstrap/D4_MSC_bootstrap_<i>.parquet`
rather than returned in memory.

```{r bootstrap, eval = FALSE}
build_study_cohort(
    # ... same arguments as above ...
    bootstrap_pars = list(
        with_bootstrap = TRUE,
        n_bootstraps   = 500,
        dir_bootstrap  = dir_bootstrap,
        start_seed     = 42
    )
)
```
