---
title: "Matching Without Replacement (Greedy)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Matching Without Replacement (Greedy)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    eval = FALSE
)
```

## Overview

This vignette demonstrates **greedy matching without replacement** using
`CohortBuilder`. In this mode each person can appear at most once in the final
matched cohort (as either an exposed or a control). Matching proceeds in rounds:
in each round every available exposed spell is matched to its highest-priority
available control, and all matched persons are then removed from both pools
before the next round begins.

This mode does not currently support bootstrap resampling. Use
`matching_mode = "with_replacement"` (see the companion vignette) for
bootstrap-based inference.

## Prerequisites

```{r setup}
library(CohortBuilder)
library(data.table)
```

## Step 1 — Configure parameters

```{r parameters}
# Variables to match on (must be present in the eligibility data)
matching_vars <- c(
    "SV_SEX", "SV_REGION", "SV_HIST_COVID_VACC", "SV_PRIOR_COVID_DG",
    "SV_BRAND_COVID_VACC", "SV_PREG_STATUS", "SV_IMMUNOCOMPROMISED",
    "CDC_RISK", "SV_SES_STATUS"
)

# --- Directories --------------------------------------------------------------
dir_input <- "input"
dir_int_outputs <- "intermediate_outputs"
dir_matching_db <- "intermediate_outputs/matching.duckdb"
dir_bootstrap <- "intermediate_outputs/bootstraps"
dir_outputs <- "outputs"

# --- Matching options ----------------------------------------------------------
save_output <- FALSE
save_intermediate_matching_ouputs <- TRUE
with_bootstrap <- FALSE
n_bootstraps <- 1
start_seed <- 42
```

## Step 2 — Prepare input data

Provide your own eligibility dataset or generate synthetic data for testing:

```{r input}
# Option A: generate synthetic data
D3_ELIGIBILITY <- generate_eligibility_data(
    n           = 100000,
    start_seed  = start_seed,
    save_output = FALSE,
    output_dir  = dir_input,
    output_file = "D3_ELIGIBILITY"
)

# Option B: read your own data
# D3_ELIGIBILITY <- arrow::read_parquet(file.path(dir_input, "D3_ELIGIBILITY.parquet"))
```

The minimum required columns are the same as for the with-replacement mode:

| Column | Description |
|---|---|
| `person_id` | Unique person or spell identifier |
| Matching variables | As listed in `matching_vars` |
| `eligible_exposed` | Logical — TRUE if the row is an exposed spell |
| `eligible_control` | Logical — TRUE if the row is an eligible control |
| `matching_status_start` | Date: start of the matching window |
| `matching_status_end` | Date: end of the matching window |
| `year_of_birth` | Integer year of birth (used for ± 1-year age matching) |

## Step 3 — Load SQL templates

The no-replacement query uses a `__START_SEED__` placeholder that is replaced at
runtime with `start_seed` to ensure deterministic control selection:

```{r sql}
matching_query <- suppressWarnings(
    getSQL(system.file("sql_queries", "matching_query_without_replacement.sql",
        package = "CohortBuilder"
    ))
)
```

## Step 4 — Run the pipeline

```{r run}
D4_MSC <- build_study_cohort(
    eligible_pop = D3_ELIGIBILITY,
    matching_query = matching_query,
    matching_vars = matching_vars,
    dir_matching_db = dir_matching_db,
    n_cores = NULL, # auto-detects all cores minus one
    log_name = "log_build_study_cohort",
    output_pars = list(
        save_output      = save_output,
        dir_output       = dir_outputs,
        output_file_name = "D4_MSC"
    ),
    intermediate_output_pars = list(
        save_intermediate_outputs = save_intermediate_matching_ouputs,
        dir_intermediate_outputs  = dir_int_outputs,
        profile_table_name        = "D3_LOOKUP_TABLE",
        matching_pop_name         = "D3_MATCHING_POP"
    ),
    bootstrap_pars = list(
        with_bootstrap = with_bootstrap,
        n_bootstraps   = n_bootstraps,
        dir_bootstrap  = dir_bootstrap,
        start_seed     = start_seed
    ),
    input_column_names = list(
        col_person_id             = "person_id",
        col_eligible_exposed      = "eligible_exposed",
        col_eligible_control      = "eligible_control",
        col_matching_status_start = "matching_status_start",
        col_matching_status_end   = "matching_status_end",
        col_age_iterator          = "year_of_birth"
    ),
    output_column_names = list(
        col_person_id       = "person_id",
        col_match_id        = "match_id",
        col_treatment_group = "group",
        col_T0              = "T0"
    ),
    matching_mode = "without_replacement"
)
```

## Step 5 — Inspect results

`D4_MSC` is a `data.table` with one row per person-spell:

```{r inspect}
# Group counts — every exposed spell is accounted for
D4_MSC[, .N, by = group]
#>       group      N
#> 1:  EXPOSED  xxxxx
#> 2:  CONTROL  xxxxx
#> 3: UNMATCHED xxxxx

# Verify strict no-replacement: each person appears at most once among matched rows
matched <- D4_MSC[!is.na(match_id)]
matched[, .N, by = person_id][N > 1] # should be empty
```

| Column | Description |
|---|---|
| `person_id` | Person/spell identifier |
| `match_id` | Shared identifier for each matched pair; `NA` for unmatched exposed |
| `boot_id` | Always 1 in no-replacement mode |
| `group` | `"EXPOSED"`, `"CONTROL"`, or `"UNMATCHED"` |
| `T0` | Index date — start of the exposed spell (shared within each pair) |
| `matching_status_start` | Start of the eligibility window |
| `matching_status_end` | End of the eligibility window |
| Matching variables | Copied from the lookup table |

## Algorithm details

The greedy algorithm works as follows:

1. **Priority ordering** — All available exposed spells are ranked by ascending
   `matching_status_start`, with seeded hash-based tie-breaking so results are
   fully reproducible.

2. **Candidate generation** — For each exposed spell, control candidates are
   filtered to those that share the same `groupkey` (matching profile), have an
   overlapping eligibility window, have a different `person_id`, and have a
   year of birth within ± 1 of the exposed spell.

3. **Control selection** — Among the candidates for a given exposed spell, the
   single highest-priority control is selected using the same seeded hash, then
   de-duplicated so each control can be selected by at most one exposed spell
   per round.

4. **Person-level collision resolution** — Within a round, a person cannot
   appear as both an exposed and a control. Where a collision would occur, the
   lower-priority match is dropped.

5. **Pool update** — All matched persons are removed from both pools and the
   next round begins.

6. **Termination** — The algorithm stops when no new matched pairs are produced
   in a round.

Exposed spells that are never matched are returned with `group = "UNMATCHED"`
and `match_id = NA`.
